---
title: "Exploratory Data Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
dat = read.csv('brfss_final.csv')
```
### Explore target variable
```{r}
MI_dat <- dat
MI_dat$CVDINFR4 <- ifelse(MI_dat$CVDINFR4 > 2, 3, MI_dat$CVDINFR4)
ggplot(MI_dat, aes(x = factor(CVDINFR4))) +
  geom_bar(aes(fill = factor(CVDINFR4))) +
  scale_fill_manual(values = c("#f4a599", "#92c5de", "grey"), name = "Ever Reported Having MI") +
  labs(title = "Myocardial Infarction Bar Plot", x = "Myocardial Infarction", y = "Count") +
  scale_x_discrete(labels = c("Yes", "No", "Missing")) +
  theme(legend.position = "none")
```


### Variable Importance
```{r}
library(vip)
library(glmnet)
#define outcome variable
y <- dat[,"MICHD"] |> as.matrix()
#define matrix of predictor variables
x <- dat[, !names(dat) %in% c("MICHD", "CVDINFR4","CVDCRHD4")] |> as.matrix()
elasticnet.mod = glmnet(x,y,alpha=0.5,family="binomial")
vip(elasticnet.mod, num_features=10, geom = "point")
```

#### Explore useful features

Sex 
```{r}
dat %>% 
  select(SEX,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(SEX)),position="fill") +
  scale_fill_manual("Sex", values=c("#f4a599", "#92c5de"), 
                    labels=c("Male", "Female")) +
  labs(title="Sex vs MI/CHD",
       x="MI/CHD",
       y="Sex",
       fill="Gender") +
  scale_x_discrete(labels = c("Yes", "No"))
```

DIABETE4: (Ever told) (you had) diabetes?
```{r}
dat_db <- dat
dat_db$DIABETE4 <- ifelse(dat_db$DIABETE4 > 4, 5, dat_db$DIABETE4)
dat_db$DIABETE4 <- ifelse(dat_db$DIABETE4 == 2, 1, dat_db$DIABETE4)
dat_db$DIABETE4 <- ifelse(dat_db$DIABETE4 == 4, 3, dat_db$DIABETE4)
dat_db %>% select(DIABETE4,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(DIABETE4)),position="fill") +
  scale_fill_brewer("Diabetes History", palette="Reds",labels = c("Yes", "No", "No Response")) +
  labs(title="Diabetes vs MI/CHD",
       x="MI/CHD",
       y="Diabetes") +
  scale_x_discrete(labels = c("Yes", "No"))
```

CHCCOPD3: (Ever told) (you had) C.O.P.D. (chronic obstructive pulmonary disease), emphysema or chronic bronchitis?
```{r}
dat_copd <- dat
dat_copd$CHCCOPD3 <- ifelse(dat_copd$CHCCOPD3  > 2, 3, dat_copd$CHCCOPD3)
dat_copd %>% select(CHCCOPD3,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(CHCCOPD3)),position="fill") +
  scale_fill_brewer("COPD History", palette="Reds",labels = c("Yes", "No", "No Response")) +
  labs(title="COPD vs MI/CHD",
       x="MI/CHD",
       y="COPD") +
  scale_x_discrete(labels = c("Yes", "No"))
```

RFHYPE6: Adults who have been told they have high blood pressure by a doctor, nurse, or other health professional [1 = NO, 2 = YES]
```{r}
dat_hb <- dat
dat_hb$RFHYPE6 <- ifelse(dat_hb$RFHYPE6  > 2, 3, dat_hb$RFHYPE6)
dat_hb %>% select(RFHYPE6,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(RFHYPE6)),position="fill") +
  scale_fill_brewer("High Blood Pressure History", palette="Reds",
                    labels = c("No", "Yes", "No Response")) +
  labs(title="High Blood Pressure vs MI/CHD",
       x="MI/CHD",
       y="High Blood Pressure") +
  scale_x_discrete(labels = c("No", "Yes"))
```

SMOKER3: Four-level smoker status: Everyday smoker, Someday smoker, Former smoker, Non-smoker [1-4]
```{r}
dat %>% select(SMOKER3,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(SMOKER3)),position="fill") +
  scale_fill_brewer("Smoking Status", palette="Reds",
                    labels = c(" Everyday smoker", "Someday smoker", "Former smoker",
                               "Non-smoker", "No Response")) +
  labs(title="Smoking vs MI/CHD",
       x="MI/CHD",
       y="Smoking") +
  scale_x_discrete(labels = c("No", "Yes"))

##### although not in variable importance plot, explore continuous variable about smoking#######

#LCSNUMCG: On average, when you smoke/smoked regularly, about how many cigarettes do/did you usually smoke each day?
ggplot(dat, aes(x = LCSNUMCG, y = MICHD, color = factor(SEX))) +
  stat_smooth(method = "loess", alpha = 0.1, size = 2, aes(fill=factor(SEX))) +
  labs(x = "Number of Cigarettes Per Day", y = "MICHD")

library(reshape2)
dat_heatmap <- dat %>% select(LCSNUMCG, MICHD) %>% drop_na()
# Bin the continuous predictor into groups
dat_heatmap$LCSNUMCG_group <- cut(dat_heatmap$LCSNUMCG, breaks = 10)
heatmap_data <- dat_heatmap %>% group_by(LCSNUMCG_group, MICHD) %>% summarize(count = n())
heatmap_data_wide <- dcast(heatmap_data, LCSNUMCG_group ~ MICHD, value.var = "count")
my_palette <- colorRampPalette(c("white", "red"))



#LCSFIRST: How old were you when you first started to smoke cigarettes regularly



```



General Health
```{r}
dat %>% select(GENHLTH,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(GENHLTH)),position="fill") +
  scale_fill_brewer("General Health Conditions", palette="Reds") +
  labs(title="General Health Conditions vs MI/CHD",
       x="MI/CHD",
       y="Categories of General Health") +
  scale_x_discrete(labels = c("Yes", "No"))
```


QSTVER: Questionnaire Version Identifier
10 Landline single questionnaire
11 Landline multiple questionnaire version 1
12 Landline multiple questionnaire version 2
13 Landline multiple questionnaire version 3
```{r}
dat %>% select(QSTVER,MICHD) %>%
  drop_na() %>%
  ggplot() +
  geom_bar(aes(x=factor(MICHD),fill=factor(QSTVER)),position="fill") +
  scale_fill_brewer(" Questionnaire Version", palette="Reds") +
  labs(title=" Questionnaire Version vs MI/CHD",
       x="MI/CHD",
       y=" Questionnaire Version")
```




May be useful someday...
```{r}
dat %>% select(GENHLTH,QSTLANG,MICHD) %>%
  drop_na() %>%
  group_by(factor(GENHLTH),factor(QSTLANG)) %>%
  ggplot() +
    geom_col(aes(x=factor(GENHLTH),y=factor(MICHD),fill=factor(QSTLANG)),
             position="dodge") +
    #theme(axis.text.x = element_text(angle=90, vjust = 0.5,hjust=1)) + 
    labs(title="MI/CHD vs General health",
         x="General Health",
         y="MI/CHD") + 
    scale_fill_manual(values=c('orange','grey50'))
```





